\documentclass[12pt, a4paper]{article}

% --- Font and Language Setup (Requires XeLaTeX compiler) ---
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{carlito}
\renewcommand{\familydefault}{\sfdefault}

% --- Packages ---
\usepackage{geometry}
\geometry{top=2.5cm, bottom=2.5cm, left=2.5cm, right=2.5cm} % Standard academic margins
\usepackage{graphicx}
\usepackage{titlesec}
\usepackage[hidelinks]{hyperref}
\usepackage{xcolor}
\usepackage{listings}

% --- Python Code Formatting Setup ---
\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}
\lstset{style=mystyle}
\renewcommand{\lstlistingname}{Code Snippet}

% --- Document Info ---
\title{
    \Large \textbf{ITNPAI2: AI for Computer Vision} \\
    \vspace{0.5cm}
    \LARGE \textbf{Ariadne} \\
    \vspace{0.25cm}
    \large A Synthetic Proof of Concept for Air-to-Ground Optical Navigation
}
\author{Student ID: 3539054}
\date{April 1, 2026}

\begin{document}

\maketitle

\section{Problem Statement and Contextualisation}

\quad In classical mythology, the labyrinth of Crete was navigated with \textit{"Ariadne’s thread,"} a lifeline provided by an outside observer to guide a vulnerable explorer safely through a complex, unknown maze. Modern autonomous robotics faces a similar problem. Unmanned Ground Vehicles (UGVs) used in search and rescue (SAR) or tactical situations typically rely on Global Positioning Systems (GPS) or active sensor emissions, such as LiDAR, to determine their location and plan navigable routes. \cite{11219502}. However, in highly contested or unstructured environments, such as disaster zones, deep forest canopies, or electronic warfare scenarios, these signals can often be degraded, obstructed, or intentionally jammed \cite{8741612}.

\vspace{0.5cm}

When active signals fail, autonomous ground navigation systems become functionally blind. To address this critical vulnerability, this project presents \textbf{Ariadne}: an autonomous \textbf{air-to-ground optical routing pipeline}. This system acts as a modern digital thread, utilising a single high-altitude RGB image captured by an Unmanned Aerial Vehicle (UAV). It passively maps the environment below, identifies navigable corridors, and calculates the optimal path for the ground asset.

\vspace{0.5cm}

To test the core logic of the computer vision and pathfinding modules without the interference of unpredictable lighting conditions or organic textures found in the real world, this study uses procedurally generated mazes as a synthetic proof-of-concept. The maze environment acts as a controlled sandbox to isolate and evaluate the core algorithmic pipeline before scaling up to more complex, unstructured terrains.

\vspace{0.5cm}

\subsection{System Requirements}

To successfully achieve optical-based autonomous routing, the proposed pipeline must fulfil the following sequential requirements:

\begin{enumerate}
    \item \textbf{Target Extraction:} The system must accurately detect and extract the pixel coordinates of the starting agent (the UGV) and the target destination (a human or objective) from the overhead image.

    \item \textbf{Environmental Segmentation:} The system must classify the environment, accurately dividing the image into strictly traversable (pathways) and non-traversable (walls/obstacles) terrain.

    \item \textbf{Tactical Pathfinding:} The system must compute an optimal, collision-free route from the start node to the target node using the segmented map, and project this route back onto the original visual space.
\end{enumerate}

\vspace{0.5cm}

\section{Literature Review}

The difficulty of autonomous navigation in environments without GPS has led to significant research into alternative localisation and mapping methods. 

\vspace{0.5cm}

Recent research on Unmanned Ground Vehicles (UGVs) has extensively explored the use of environment-specific, non-visual sensors to overcome the limitations posed by degraded optical conditions. The project applied \textbf{magnetic field map-based localisation}, which uses natural magnetic anomalies as navigational features \cite{11219502}. This approach employs Gaussian Process (GP) Regression to create high-resolution magnetic signature maps. It also combines a hybrid Particle Filter (PF) and Unscented Kalman Filter (UKF) to estimate the trajectory of UGVs. In tests conducted in indoor corridors, this methodology demonstrated remarkable resilience, reducing localisation errors by as much as 45\% compared to standard particle filter (PF)-based methods. However, this approach relies entirely on localised, ground-level sensing, meaning that the UGV must physically navigate and explore the environment to use the map effectively. \textbf{Ariadne} addresses this critical gap by changing the approach from \textit{"ground-level discovery"} to \textit{"overhead deterministic routing,"} enabling the UGV to receive an optimal path before encountering any obstacles.

\vspace{0.5cm}

Beyond ground-level localisation, extensive research has examined how Unmanned Aerial Vehicles (UAVs) navigate in GPS-denied environments. A prominent framework proposed for UAV exploration combines Simultaneous Localisation and Mapping (\textbf{SLAM}) with Partially Observable Markov Decision Processes (\textbf{POMDP}) \cite{8741612}. In this model, navigation is treated as a sequential decision problem under uncertainty. Simulation results showed that the UAV could safely navigate, avoid collisions, and maximise information gain while creating an occupancy map of an unknown environment. While effective for a single agent, this framework focuses solely on the UAV’s self-preservation and exploration. It does not consider the potential of utilising the generated occupancy map to actively guide a separate ground asset. \textbf{Ariadne} bridges this gap by using the UAV as an overhead optical \textit{"oracle"} that segments the terrain to guide the movements of a UGV.

\vspace{0.5cm}

To achieve optical terrain mapping, deep learning-based semantic segmentation has become the standard. Recent advancements in aerial road extraction utilise Convolutional Neural Networks (CNNs), particularly the \textbf{U-Net architecture}, to carry out binary semantic segmentation on remote sensing imagery \cite{Doghmane_Boukari_2024}. Research by Doghmane and Boukari demonstrates that enhancing the U-Net model with \textbf{Batch Group Normalisation} (BGN) significantly improves its ability to classify individual pixels into distinct object categories, such as road versus non-road areas. By dividing high-resolution aerial images into manageable patches and applying a sigmoid activation function to output a continuous pixel value, their BGN-UNet model achieved an exceptional Mean Intersection over Union (IoU) of 98.4\%. \textbf{Ariadne} takes direct inspiration from this pixel-wise classification approach. Just as BGN-UNet isolates navigable infrastructure from complex backgrounds, Ariadne utilises deep learning segmentation on overhead UAV imagery to divide simulated tactical environments (represented via procedural mazes) into strictly traversable paths and non-traversable boundaries. However, while current literature primarily treats the segmented image as the final output, Ariadne expands on this idea by utilising the high-IoU segmentation mask as a foundational grid for \textbf{A* pathfinding}. This approach translates the visual classification into a mathematical route for a ground agent.


\bibliographystyle{ieeetr}
\bibliography{references}
    
\end{document}
